---
title: "BiteAnalysisPreliminary"
output: html_document
---

install.packages("anytime")
install.packages("hms")
install.packages("lessR")
library(anytime)
library(hms)
library(lubridate)
library(tidyverse)
library(fuzzyjoin)
library(lme4)
library(emmeans)
library(lessR)

```

## 

# This chunk reads in .csvs that contain bite information, focal pair ID, group, phenotype, dominance, reproductive status, weather data, fruit biomass data, and behavior duration data

```{r}
df_bite <- read.csv("bite_query_correct_filtered.csv")
df_pair <- read.csv("pair_ID.csv")
df_group <- read.csv("group_iD.csv")
df_phenotype <- read.csv("focal_phenotypes.csv")
df_dom <- read.csv("dom_2019.csv")
df_rs <- read.csv("reproductivestatus_2019.csv")
df_weather <- read.csv("Weather_For_Allegra_2019.csv")
df_biomass <- read.csv("monthly_fruit_biomass.csv")
df_duration <- read.csv("Foraging_State_Duration.csv")

```

#This chunk cleans up the data frames from above; removes/renames columns, sets Date objects, etc.

``` {r}

#Removing DateTime column from the dominance df
df_dom <- df_dom %>%
                      dplyr::select(-DateTime) 


#Renaming month column in df_biomass
df_biomass <- df_biomass %>% 
           rename(
           Month = ï..Month,
   
                         )


#Setting start data and end data in the reproductive status dataframe as date objects
df_rs$StartDate <- parse_date_time(df_rs$StartDate, orders = "mdy")
df_rs <- df_rs %>% mutate(StartDate = as_datetime(StartDate))
df_rs$StartDate <- as.Date(df_rs$StartDate)
df_rs$EndDate <- parse_date_time(df_rs$EndDate, orders = "mdy")
df_rs <- df_rs %>% mutate(EndDate = as_datetime(EndDate))
df_rs$EndDate <- as.Date(df_rs$EndDate)

#Renaming DateTime
df_weather <- df_weather %>% 
           rename(
           DateTime = ï..DateTime,
   
                         )

#Setting DateTime as a date object in df_weather
df_weather$DateTime <- parse_date_time(df_weather$DateTime, orders = "ymd")
df_weather <- df_weather %>% mutate(DateTime = as_datetime(DateTime))
df_weather$DateTime <- as.Date(df_weather$DateTime)

#Changing name of DateTimeBegin to DateTime in bite dataframe
df_bite <- df_bite %>% 
           rename(
           DateTime = DateTimeBegin,
   
                         )


#Setting date time in the bite dataframe as date object
df_bite$DateTime <- parse_date_time(df_bite$DateTime, orders = "mdy")
df_bite <- df_bite %>% mutate(DateTime = as_datetime(DateTime))
df_bite$DateTime <- as.Date(df_bite$DateTime)



#Renaming focal column
df_bite <- df_bite %>% 
           rename(
           Focal = ï..ID,
   
                         )
#Renaming pair column                         
df_pair <- df_pair %>% 
           rename(
           Pair = ï..Pair,
   
                         )
                         
#Renaming animal column
df_phenotype <- df_phenotype %>% 
           rename(
           Animal = ï..Animal,
   
                         )    

#Renaming duration column
df_duration <- df_duration %>% 
           rename(
           Focal = ï..ID,
   
                         )
                         
#Excluding Ducky, setting as data frame, and dropping unused factor levels for df_duration

df_duration <- df_duration[!(df_duration$Animal == "DU") ,]
df_duration <- df_duration[!(df_duration$Focal  == "226") ,]
df_duration<- as.data.frame(df_duration)
df_duration <- droplevels(df_duration)                        
                         
                         
#Subsetting durations to include only foraging
df_foraging_duration <- df_duration[df_duration$State == "FFR" | df_duration$State == "FIN" | df_duration$State == "EFI" | df_duration$State == "FOT" | df_duration$State == "FUN" | df_duration$State == "EFF" | df_duration$State == "FFL", ]                         
                      
                         
```

```{r}

#Setting duration as seconds
df_foraging_duration$Duration <- lubridate::hms(df_foraging_duration$Duration)
df_foraging_duration$Duration <- as.numeric(df_foraging_duration$Duration)


#Summing duration for every foraging state per focal, and storing as new variable, "Duration"
df_grouped_foraging_state <- df_foraging_duration %>% 
  group_by(Focal, Animal) %>%
  summarise(Duration = sum(Duration)) 


#Grouping by variables Focal, Animal, State, and Behavior
df_grouped <- df_bite %>% 
  group_by(Focal, DateTime, Animal, State, BehaviorCode) 

#Counting bites per foraging state for each focal
df_grouped2 <- df_grouped %>%  summarise(Count=n())

#Making bites per foraging state their own column
df_grouped3 <- df_grouped2 %>% spread(State,Count)

#Converting NAs to zeroes
df_grouped3[c("EFF", "EFI", "FFL", "FFR", "FIN","FOT","FUN")][is.na(df_grouped3[c("EFF", "EFI", "FFL", "FFR", "FIN","FOT","FUN")])] <- 0

#Excluding Ducky, for whom there is one focal
df_grouped3 <- df_grouped3[!(df_grouped3$Animal == "DU") ,]
df_grouped3 <- df_grouped3[!(df_grouped3$Focal  == "226") ,]

#Setting as data frame and dropping unused levels
df_grouped3 <- as.data.frame(df_grouped3)
df_grouped3 <- droplevels(df_grouped3)

  
#Joining duration column to bite data frame
df_duration_bite <- df_grouped3 %>% left_join(., df_grouped_foraging_state, by = c("Focal" = "Focal", 
                                                            "Animal" = "Animal"))
#Joining focal pair id to bite data frame                                                            
df_duration_bite <- df_pair %>% left_join(., df_duration_bite, by = c( 
                                                            "Animal" = "Animal"))                                   
#Joining group info to bite data frame
df_duration_bite <- df_group %>% left_join(., df_duration_bite, by = c( 
                                                            "Animal" = "Animal"))               
#Joining color vision phenotype data to bite dataframe                                  
df_duration_bite <- df_phenotype %>% left_join(., df_duration_bite, by = c( 
                                                            "Animal" = "Animal")) 
                                                            
#Joining dominance data to bite dataframe                                                           
df_duration_bite <- df_dom %>% left_join(., df_duration_bite, by = c( 
                                                            "Animal" = "Animal"))    

                                                       
                                                        
#Converting dominance ranks to ordinal variable with levels low, mid, and high                                                       
df_duration_bite$Scaled.Rank <- if_else(df_duration_bite$Scaled.Rank <= 0.33, "low", if_else(df_duration_bite$Scaled.Rank > 0.33 & df_duration_bite$Scaled.Rank <= 0.66, "mid","high"))                                                            
      
                                                            
#Fuzzy left joining reproductive status to behavioral data
final_df <- df_duration_bite %>%
            fuzzy_left_join(., df_rs, by = c("Animal" = "Animal", "DateTime" = "StartDate", "DateTime" = "EndDate"), match_fun = list(`==`, `>=`, `<=`))            
            
 
#Getting rid of extra columns from the fuzzy left join
 final_df <- final_df %>%
                      dplyr::select(-Animal.y, -StartDate, -EndDate)            
            
#Renaming animal
final_df <- final_df %>%
              rename(Animal = Animal.x)                                             

#Renaming scaled rank        
final_df <- final_df %>%
              rename(DominanceCategory = Scaled.Rank)

#Renaming reproductive status
final_df <- final_df %>%
              rename(ReproductiveStatus = Reproductive.Status)
              
#Removing one focal where the tablet wasn't recording seconds, causing a 0 duration, which messes up the offset in the models

final_df <- final_df %>% 
            filter(Duration > 0)
            
#Joining weather variables to final df
final_df <- final_df %>% left_join(., df_weather, by = c( 
                                                         "DateTime" = "DateTime")) 
                                                         
#Adding month column to final_df to facilitate join with df_biomass
final_df <- final_df %>% mutate(Month = month(DateTime))
final_df <- final_df %>% left_join(., df_biomass, by = c( 
                                                          "Month" = "Month"))  
                                                          

#Renaming monthly fruit biomass
final_df <- final_df %>%
              rename(MonthlyFruitBiomass = Ripe.Fruit.Biomass)

#Recoding D and T (this recode function is from the package car)  
final_df$Phenotype <- Recode(final_df$Phenotype, "'D'='Dichromat'; 'T'='Trichromat'")

#Z-transforming TempMax, TempMin, Rainfall, and Monthly Fruit Biomass
final_df_transformed <- final_df %>%
       mutate_at(vars("TempMax", "TempMin", "Rain", "MonthlyFruitBiomass"), list(~as.numeric(scale(.)))) 
       
last_df <- final_df_transformed %>% mutate(FruitInsectSum = FFR + FIN)      
last_df <- last_df %>% mutate(FruitProportion = FFR / FruitInsectSum)        
last_df <- last_df %>% mutate(InsectProportion = FIN / FruitInsectSum) 
```

#GLMM examining the effect of phenotype, dominance, and reproductive status on bites of fruit per focal, controlling for animal, group, and pair ID

FruitBiteModel <- glmer(FFR ~  
                                Phenotype + 
                                DominanceCategory +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = final_df_transformed,
                              family = poisson(link = "log"))          

drop1(FruitBiteModel, test = "Chisq")



InsectBiteModel <- glmer(FIN ~  
                                Phenotype + 
                                DominanceCategory +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = final_df_transformed,
                              family = poisson(link = "log"))          

drop1(InsectBiteModel, test = "Chisq")




InsectBiteModel.emm <- emmeans(InsectBiteModel, "Phenotype", type = "response", adjust = "tukey") #This function displays the LSM

pairs(emmeans(InsectBiteModel, "Phenotype", adjust = "tukey"))



InsectBite.emm.cld.tukey <- cld.emmGrid(InsectBiteModel.emm, alpha = 0.05, Letters = letters, adjust = "tukey") #This function creates a compact letter display of the LSM enabling use with ggplot. It also adds a Tukey adjustment to test for honest significant difference between the means.

```
Plot
```{r}
InsectBite <- InsectBite.emm.cld.tukey #Renaming CLD object for increased ease with plotting.

InsectBitePlot <- ggplot(InsectBite, aes(x = Phenotype, y = rate, shape = Phenotype, color = Phenotype)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("D" = 16, "T" = 16)) +
  scale_color_manual(values = c("D" = "#0072B2", "T" = "#E69F00")) +
  geom_errorbar(aes(ymin  =  rate - SE, ymax  =  rate + SE), width =  0.2, size  =  0.7) +
  scale_y_continuous(limits = c(20,60)) +
  scale_x_discrete(limits = c("D", "T")) +
  ggtitle("Insect Bites per Focal") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.y = element_text(size = 15, color = "black")) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(legend.position = "bottom")

InsectBitePlot

```



InsectProportionBiteModel <- glmer(InsectProportion ~  
                                Phenotype + 
                                ScaledRank +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                #offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = last_df,
                              family = binomial(link = "logit"))
                              
drop1(InsectProportionBiteModel, test = "Chisq")


FruitProportionBiteModel <- glmer(FruitProportion ~  
                                Phenotype + 
                                ScaledRank +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                #offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = last_df,
                              family = binomial(link = "logit"))
                              
drop1(FruitProportionBiteModel, test = "Chisq")


```{r}

#Seperately analyzing bites of conspicuous fruits

df_bite_fruit <- df_bite[df_bite$State == "FFR", ] 

df_color <- read.csv("colors_final_final.csv")
df_bite_fruit <- df_bite_fruit %>% left_join(., df_color, by = c("Taxon" = "Taxon" 
))

df_bite_fruit <- df_bite_fruit[df_bite_fruit$ColourCategory == "con" |df_bite_fruit$ColourCategory == "cry" | df_bite_fruit$ColourCategory == "dark"| df_bite_fruit$ColourCategory == "con_both",]


#Grouping by variables Focal, Animal, State, and Behavior
dfz <- df_bite_fruit %>% 
  group_by(Focal, DateTime, Animal, ColourCategory) 

#Counting bites per foraging state for each focal
dfz1 <- dfz %>%  summarise(Count=n())

#Making bites per foraging state their own column
dfz2 <- dfz1 %>% spread(ColourCategory,Count)

#Removing the NA column
dfz2 <- dfz2 %>%
              dplyr::select(- "<NA>")


#Setting NAs to zero
dfz2[c("con", "cry", "dark", "con_both")][is.na(dfz2[c("con", "cry", "dark", "con_both")])] <- 0


#Joining duration column to conspicuous fruit bite data frame
df_con_bite_duration <- dfz2 %>% left_join(., df_grouped_foraging_state, by = c("Focal" = "Focal", 
                                                            "Animal" = "Animal"))
#Joining focal pair id to bite data frame                                                            
df_con_bite_duration <- df_pair %>% left_join(., df_con_bite_duration, by = c( 
                                                            "Animal" = "Animal"))                                   
#Joining group info to bite data frame
df_con_bite_duration <- df_group %>% left_join(., df_con_bite_duration, by = c( 
                                                            "Animal" = "Animal"))               
#Joining color vision phenotype data to bite dataframe                                  
df_con_bite_duration <- df_phenotype %>% left_join(., df_con_bite_duration, by = c( 
                                                            "Animal" = "Animal")) 
                                                            
#Joining dominance data to bite dataframe                                                           
df_con_bite_duration <- df_dom %>% left_join(., df_con_bite_duration, by = c( 
                                                            "Animal" = "Animal"))    

                                                       
                                                        
#Converting dominance ranks to ordinal variable with levels low, mid, and high                                                       
df_con_bite_duration$Scaled.Rank <- if_else(df_con_bite_duration$Scaled.Rank <= 0.33, "low", if_else(df_con_bite_duration$Scaled.Rank > 0.33 & df_con_bite_duration$Scaled.Rank <= 0.66, "mid","high"))                                                            
      
                                                            
#Fuzzy left joining reproductive status to behavioral data
df_con_bite_duration <- df_con_bite_duration %>%
            fuzzy_left_join(., df_rs, by = c("Animal" = "Animal", "DateTime" = "StartDate", "DateTime" = "EndDate"), match_fun = list(`==`, `>=`, `<=`))            
            
 
#Getting rid of extra columns from the fuzzy left join
df_focal_cat_final <- df_con_bite_duration  %>%
                      dplyr::select(-Animal.y, -StartDate, -EndDate)            
            
#Renaming animal
df_focal_cat_final <- df_focal_cat_final %>%
              rename(Animal = Animal.x)                                             

#Renaming scaled rank        
df_focal_cat_final <- df_focal_cat_final %>%
              rename(DominanceCategory = Scaled.Rank)

#Renaming reproductive status
df_focal_cat_final <- df_focal_cat_final %>%
              rename(ReproductiveStatus = Reproductive.Status)
              
            
#Joining weather variables to final df
df_focal_cat_final <- df_focal_cat_final %>% left_join(., df_weather, by = c( 
                                                         "DateTime" = "DateTime")) 
                                                         
#Adding month column to final_df to facilitate join with df_biomass
df_focal_cat_final <- df_focal_cat_final %>% mutate(Month = month(DateTime))
df_focal_cat_final <- df_focal_cat_final %>% left_join(., df_biomass, by = c( 
                                                          "Month" = "Month"))  
                                                          

#Renaming monthly fruit biomass
df_focal_cat_final <- df_focal_cat_final %>%
              rename(MonthlyFruitBiomass = Ripe.Fruit.Biomass)

df_focal_cat_final <- df_focal_cat_final %>%
       mutate_at(vars("TempMax", "TempMin", "Rain", "MonthlyFruitBiomass"), list(~as.numeric(scale(.)))) 


ConFruitBiteModel <- glmer(con ~  
                                Phenotype + 
                                DominanceCategory +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = df_focal_cat_final,
                              family = poisson(link = "log"))          

drop1(ConFruitBiteModel, test = "Chisq")

ConFruitBiteModel.emm <- emmeans(ConFruitBiteModel, "Phenotype", type = "response", adjust = "tukey") #This function displays the LSM


pairs(emmeans(ConFruitBiteModel, "Phenotype", adjust = "tukey"))




ConFruitBiteModel.emm.cld.tukey <- cld.emmGrid(ConFruitBiteModel.emm, alpha = 0.05, Letters = letters, adjust = "tukey") #This function creates a compact letter display of the LSM enabling use with ggplot. It also adds a Tukey adjustment to test for honest significant difference between the means.

```
Plot
```{r}
ConFruitBite <- ConFruitBiteModel.emm.cld.tukey #Renaming CLD object for increased ease with plotting.

ConFruitBitePlot <- ggplot(ConFruitBite, aes(x = Phenotype, y = rate, shape = Phenotype, color = Phenotype)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("D" = 16, "T" = 16)) +
  scale_color_manual(values = c("D" = "#0072B2", "T" = "#E69F00")) +
  geom_errorbar(aes(ymin  =  rate - SE, ymax  =  rate + SE), width =  0.2, size  =  0.7) +
  scale_y_continuous(limits = c(0,70)) +
  scale_x_discrete(limits = c("D", "T")) +
  ggtitle("Conspicuous Fruit Bites per Focal") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.y = element_text(size = 15, color = "black")) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  theme(legend.position = "bottom")

ConFruitBitePlot




```






















ExtractiveInsectBiteModel <- glmer(EFI ~  
                                Phenotype + 
                                ScaledRank +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = final_df_transformed,
                              family = poisson(link = "log"))          

drop1(ExtractiveInsectBiteModel, test = "Chisq")



FOTBiteModel <- glmer(FOT ~  
                                Phenotype + 
                                ScaledRank +
                                ReproductiveStatus +
                                TempMax +
                                TempMin +
                                Rain +
                                MonthlyFruitBiomass +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + (1 | Pair),
                              data = final_df_transformed,
                              family = poisson(link = "log")) 

drop1(FOTBiteModel, test = "Chisq")



