---
title: "EMM_Niche_Overlap"
author: "Allegra DePasquale"
date: "8/10/2020"
output: html_document
---

###This analysis uses the EcoSimR shell found at 
http://www.uvm.edu/~ngotelli/EcoSim/EcoSim.html

###First, I use the existing dataframes Bite_data_no_fruit and Bite_data_fruit to calculate estimated marginal means for each major resource category in each month. Most of the lines below are simply subsetting by each month and running GLMMS of each resource category in each month to get EMMs. I inputted each month's EMMs into their own Excel sheets, uploaded each of those in turn to EcoSimR, and recorded the observed Pianka index in the csv titled "Pianka_index_emm.csv" read into this script at line 668. It is this file that is read in and analyzed using Welch's t-test. 

```{r}

df_bite_no_fruit <- read.csv("Bite_Data_no_fruit.csv")
df_bite_fruit <- read.csv("Bite_Data_fruit.csv")

june <- df_bite_no_fruit[df_bite_no_fruit$Month == 6,]
june_con <- df_bite_fruit[df_bite_fruit$Month == 6,]

june_efi <- glmer(EFI ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = june,
                              family = poisson(link = "log")) 

emmeans(june_efi, pairwise ~ Phenotype, type = "response", adjust = "tukey")


june_fin <- glmer(FIN ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = june,
                              family = poisson(link = "log")) 

emmeans(june_fin, pairwise ~ Phenotype, type = "response", adjust = "tukey")

june_eff <- glmer(EFF ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = june,
                              family = poisson(link = "log")) 

emmeans(june_eff, pairwise ~ Phenotype, type = "response", adjust = "tukey")


june_ffl <- glmer(FFL ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = june,
                              family = poisson(link = "log")) 

emmeans(june_ffl, pairwise ~ Phenotype, type = "response", adjust = "tukey")


june_fot <- glmer(FOT ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = june,
                              family = poisson(link = "log")) 

emmeans(june_fot, pairwise ~ Phenotype, type = "response", adjust = "tukey")

june_con_both <- glmer(con_both ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = june_con,
                              family = poisson(link = "log")) 

emmeans(june_con_both, pairwise ~ Phenotype, type = "response", adjust = "tukey")

```

```{r}
july <- df_bite_no_fruit[df_bite_no_fruit$Month == 7,]
july_con <- df_bite_fruit[df_bite_fruit$Month == 7,]

july_efi <- glmer(EFI ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july,
                              family = poisson(link = "log")) 

emmeans(july_efi, pairwise ~ Phenotype, type = "response", adjust = "tukey")


july_fin <- glmer(FIN ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july,
                              family = poisson(link = "log")) 

emmeans(july_fin, pairwise ~ Phenotype, type = "response", adjust = "tukey")

july_eff <- glmer(EFF ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july,
                              family = poisson(link = "log")) 

emmeans(july_eff, pairwise ~ Phenotype, type = "response", adjust = "tukey")


july_ffl <- glmer(FFL ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july,
                              family = poisson(link = "log")) 

emmeans(july_ffl, pairwise ~ Phenotype, type = "response", adjust = "tukey")


july_fot <- glmer(FOT ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july,
                              family = poisson(link = "log")) 

emmeans(july_fot, pairwise ~ Phenotype, type = "response", adjust = "tukey")

july_con_emm <- glmer(con ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july_con,
                              family = poisson(link = "log")) 

emmeans(july_con_emm, pairwise ~ Phenotype, type = "response", adjust = "tukey")



july_con_both <- glmer(con_both ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july_con,
                              family = poisson(link = "log")) 

emmeans(july_con_both, pairwise ~ Phenotype, type = "response", adjust = "tukey")



july_cry <- glmer(cry ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = july_con,
                              family = poisson(link = "log")) 

emmeans(july_cry, pairwise ~ Phenotype, type = "response", adjust = "tukey")



```

```{r}
sep <- df_bite_no_fruit[df_bite_no_fruit$Month == 9,]
sep_con <- df_bite_fruit[df_bite_fruit$Month == 9,]

sep_efi <- glmer(EFI ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep,
                              family = poisson(link = "log")) 

emmeans(sep_efi, pairwise ~ Phenotype, type = "response", adjust = "tukey")


sep_fin <- glmer(FIN ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep,
                              family = poisson(link = "log")) 

emmeans(sep_fin, pairwise ~ Phenotype, type = "response", adjust = "tukey")

sep_eff <- glmer(EFF ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep,
                              family = poisson(link = "log")) 

emmeans(sep_eff, pairwise ~ Phenotype, type = "response", adjust = "tukey")


sep_ffl <- glmer(FFL ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep,
                              family = poisson(link = "log")) 

emmeans(sep_ffl, pairwise ~ Phenotype, type = "response", adjust = "tukey")


sep_fot <- glmer(FOT ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep,
                              family = poisson(link = "log")) 

emmeans(sep_fot, pairwise ~ Phenotype, type = "response", adjust = "tukey")

sep_con_emm <- glmer(con ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep_con,
                              family = poisson(link = "log")) 

emmeans(sep_con_emm, pairwise ~ Phenotype, type = "response", adjust = "tukey")



sep_con_both <- glmer(con_both ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep_con,
                              family = poisson(link = "log")) 

emmeans(sep_con_both, pairwise ~ Phenotype, type = "response", adjust = "tukey")



sep_cry <- glmer(cry ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = sep_con,
                              family = poisson(link = "log")) 

emmeans(sep_cry, pairwise ~ Phenotype, type = "response", adjust = "tukey")
```

```{r}
aug <- df_bite_no_fruit[df_bite_no_fruit$Month == 8,]
aug_con <- df_bite_fruit[df_bite_fruit$Month == 8,]

aug_efi <- glmer(EFI ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug,
                              family = poisson(link = "log")) 

emmeans(aug_efi, pairwise ~ Phenotype, type = "response", adjust = "tukey")


aug_fin <- glmer(FIN ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug,
                              family = poisson(link = "log")) 

emmeans(aug_fin, pairwise ~ Phenotype, type = "response", adjust = "tukey")

aug_eff <- glmer(EFF ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug,
                              family = poisson(link = "log")) 

emmeans(aug_eff, pairwise ~ Phenotype, type = "response", adjust = "tukey")


aug_ffl <- glmer(FFL ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug,
                              family = poisson(link = "log")) 

emmeans(aug_ffl, pairwise ~ Phenotype, type = "response", adjust = "tukey")


aug_fot <- glmer(FOT ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug,
                              family = poisson(link = "log")) 

emmeans(aug_fot, pairwise ~ Phenotype, type = "response", adjust = "tukey")

aug_con_emm <- glmer(con ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug_con,
                              family = poisson(link = "log")) 

emmeans(aug_con_emm, pairwise ~ Phenotype, type = "response", adjust = "tukey")



aug_con_both <- glmer(con_both ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug_con,
                              family = poisson(link = "log")) 

emmeans(aug_con_both, pairwise ~ Phenotype, type = "response", adjust = "tukey")



aug_cry <- glmer(cry ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = aug_con,
                              family = poisson(link = "log")) 

emmeans(aug_cry, pairwise ~ Phenotype, type = "response", adjust = "tukey")
```

```{r}
oct <- df_bite_no_fruit[df_bite_no_fruit$Month == 10,]
oct_con <- df_bite_fruit[df_bite_fruit$Month == 10,]

oct_efi <- glmer(EFI ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct,
                              family = poisson(link = "log")) 

emmeans(oct_efi, pairwise ~ Phenotype, type = "response", adjust = "tukey")


oct_fin <- glmer(FIN ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct,
                              family = poisson(link = "log")) 

emmeans(oct_fin, pairwise ~ Phenotype, type = "response", adjust = "tukey")

oct_eff <- glmer(EFF ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct,
                              family = poisson(link = "log")) 

emmeans(oct_eff, pairwise ~ Phenotype, type = "response", adjust = "tukey")


oct_ffl <- glmer(FFL ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct,
                              family = poisson(link = "log")) 

emmeans(oct_ffl, pairwise ~ Phenotype, type = "response", adjust = "tukey")


oct_fot <- glmer(FOT ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct,
                              family = poisson(link = "log")) 

emmeans(oct_fot, pairwise ~ Phenotype, type = "response", adjust = "tukey")

oct_con_emm <- glmer(con ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct_con,
                              family = poisson(link = "log")) 

emmeans(oct_con_emm, pairwise ~ Phenotype, type = "response", adjust = "tukey")



oct_con_both <- glmer(con_both ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct_con,
                              family = poisson(link = "log")) 

emmeans(oct_con_both, pairwise ~ Phenotype, type = "response", adjust = "tukey")



oct_cry <- glmer(cry ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct_con,
                              family = poisson(link = "log")) 

emmeans(oct_cry, pairwise ~ Phenotype, type = "response", adjust = "tukey")

oct_dark <- glmer(dark ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = oct_con,
                              family = poisson(link = "log")) 

emmeans(oct_dark, pairwise ~ Phenotype, type = "response", adjust = "tukey")



```

```{r}
nov <- df_bite_no_fruit[df_bite_no_fruit$Month == 11,]
nov_con <- df_bite_fruit[df_bite_fruit$Month == 11,]

nov_efi <- glmer(EFI ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov,
                              family = poisson(link = "log")) 

emmeans(nov_efi, pairwise ~ Phenotype, type = "response", adjust = "tukey")


nov_fin <- glmer(FIN ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov,
                              family = poisson(link = "log")) 

emmeans(nov_fin, pairwise ~ Phenotype, type = "response", adjust = "tukey")

nov_eff <- glmer(EFF ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov,
                              family = poisson(link = "log")) 

emmeans(nov_eff, pairwise ~ Phenotype, type = "response", adjust = "tukey")


nov_ffl <- glmer(FFL ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov,
                              family = poisson(link = "log")) 

emmeans(nov_ffl, pairwise ~ Phenotype, type = "response", adjust = "tukey")


nov_fot <- glmer(FOT ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov,
                              family = poisson(link = "log")) 

emmeans(nov_fot, pairwise ~ Phenotype, type = "response", adjust = "tukey")

nov_con_emm <- glmer(con ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov_con,
                              family = poisson(link = "log")) 

emmeans(nov_con_emm, pairwise ~ Phenotype, type = "response", adjust = "tukey")



nov_con_both <- glmer(con_both ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov_con,
                              family = poisson(link = "log")) 

emmeans(nov_con_both, pairwise ~ Phenotype, type = "response", adjust = "tukey")



nov_cry <- glmer(cry ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov_con,
                              family = poisson(link = "log")) 

emmeans(nov_cry, pairwise ~ Phenotype, type = "response", adjust = "tukey")

nov_dark <- glmer(dark ~  
                                Phenotype +
                                offset(log(Duration)) + 
                                (1 | Group/Animal) + 
                                (1 | Pair) + 
                                (1 | DominanceCategory)  +                       
                                (1 | ReproductiveStatus),
                              data = nov_con,
                              family = poisson(link = "log")) 

emmeans(nov_dark, pairwise ~ Phenotype, type = "response", adjust = "tukey")





pianka_emm <- read.csv("pianka_index_emm.csv")

pianka_plot <- ggplot(pianka_emm, aes(fill= FruitBiomassCategory, y = Pianka_Index, x = FruitBiomassCategory)) + geom_boxplot() + 
scale_fill_manual(values = wes_palette("Darjeeling2")) +
xlab(NULL) + ylab("Pianka's index") +
ggtitle("Pianka's Index in High and Low Fruit Months") +
theme(plot.title = element_text(hjust = 0.5)) +
theme_classic() 


ggsave(plot = pianka_plot, file = "pianka_plot.jpeg", width = 6, height = 4, units = "in")


t.test(pianka_emm$Pianka_Index ~ pianka_emm$FruitBiomassCategory)




```

